### 2019-11-13 paper/paper-draft.tex
11/13

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[
backend=biber,
style=chicago-authordate,
sorting=ynt
]{biblatex}

\addbibresource{bibliography.bib}
\AtEveryBibitem{%
  \clearlist{language}%
}

\title{Paper Proposal for POLSCI 688}
\author{Ben Goehring}
\date{\today}

\begin{document}
\begin{titlepage}
\maketitle
\end{titlepage}

asdfdsf \cite{hertel-fernandezBusinessAssociationsConservative2016}

\printbibliography

\end{document}